% ============================================================
%  TIDE & till — PRD & Design Document
%  Version 1.0 | May 2026
% ============================================================
\documentclass[11pt, letterpaper]{article}

% ── Packages ──────────────────────────────────────────────
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[protrusion=false,expansion=false]{microtype}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{titling}
\usepackage{fancyhdr}
\usepackage{mdframed}
\usepackage{tcolorbox}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{tikz}

\tcbuselibrary{skins, breakable}

% ── Color Palette ─────────────────────────────────────────
\definecolor{teal}      {HTML}{1A7A6E}
\definecolor{teallight} {HTML}{D4EDEA}
\definecolor{tealmid}   {HTML}{A8D8D3}
\definecolor{gold}      {HTML}{C4860A}
\definecolor{goldlight} {HTML}{FFF3CC}
\definecolor{coral}     {HTML}{C45E3E}
\definecolor{corallight}{HTML}{FDEAE3}
\definecolor{purple}    {HTML}{5B4A8A}
\definecolor{purplelight}{HTML}{ECEAF7}
\definecolor{green}     {HTML}{3A7A45}
\definecolor{greenlight}{HTML}{D6EDDB}
\definecolor{slate}     {HTML}{2C3E50}
\definecolor{slatelight}{HTML}{EBF0F5}
\definecolor{darkgray}  {HTML}{555555}
\definecolor{lightgray} {HTML}{F5F5F5}
\definecolor{midgray}   {HTML}{CCCCCC}
\definecolor{offwhite}  {HTML}{FAFAF8}
\definecolor{rowalt}    {HTML}{F0F4F3}

% ── Hyperlinks ────────────────────────────────────────────
\hypersetup{
  colorlinks=true,
  linkcolor=teal,
  urlcolor=teal,
  citecolor=teal,
  pdftitle={Tide \& Table — PRD \& Design Document},
  pdfauthor={Internal},
}

% ── Typography ────────────────────────────────────────────
\setlength{\parskip}{6pt}
\setlength{\parindent}{0pt}

% Section styling
\titleformat{\section}
  {\color{teal}\Large\bfseries}
  {\thesection.}{0.5em}{}
  [\vspace{2pt}\color{teal}\titlerule[1.5pt]\vspace{4pt}]

\titleformat{\subsection}
  {\color{slate}\large\bfseries}
  {\thesubsection}{0.5em}{}

\titleformat{\subsubsection}
  {\color{teal}\normalsize\bfseries}
  {\thesubsubsection}{0.5em}{}

\titlespacing*{\section}      {0pt}{18pt}{8pt}
\titlespacing*{\subsection}   {0pt}{14pt}{4pt}
\titlespacing*{\subsubsection}{0pt}{10pt}{2pt}

% ── Page Style ────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\color{darkgray}TIDE \& TABLE \ | \ PRD \& Design Document}
\fancyhead[R]{\small\color{teal}v1.0 --- May 2026}
\fancyfoot[L]{\small\color{darkgray}Confidential --- Internal Use Only}
\fancyfoot[R]{\small\color{teal}\thepage}
\renewcommand{\headrulewidth}{0.8pt}
\renewcommand{\headrule}{\hbox to\headwidth{\color{teal}\leaders\hrule height \headrulewidth\hfill}}
\renewcommand{\footrulewidth}{0.4pt}
\renewcommand{\footrule}{\hbox to\headwidth{\color{teal}\leaders\hrule height \footrulewidth\hfill}}

% ── Custom tcolorbox Environments ─────────────────────────

% Callout box (generic)
\tcbset{callout/.style={
  enhanced, breakable,
  colback=#1!8!white,
  colframe=#1,
  boxrule=0.6pt,
  left=4pt, right=4pt, top=4pt, bottom=4pt,
  fonttitle=\bfseries\small\color{#1},
  attach boxed title to top left={yshift=-2mm, xshift=4mm},
  boxed title style={colback=#1!8!white, colframe=#1, boxrule=0.4pt},
  beforeafter skip=8pt,
}}

\newtcolorbox{tealbox}[1][]{
  enhanced, breakable,
  colback=teallight, colframe=teal,
  leftrule=4pt, toprule=0.6pt, bottomrule=0.6pt, rightrule=0.6pt,
  left=8pt, right=6pt, top=5pt, bottom=5pt,
  fonttitle=\bfseries\small\color{teal},
  title={#1},
  beforeafter skip=8pt,
}

\newtcolorbox{goldbox}[1][]{
  enhanced, breakable,
  colback=goldlight, colframe=gold,
  leftrule=4pt, toprule=0.6pt, bottomrule=0.6pt, rightrule=0.6pt,
  left=8pt, right=6pt, top=5pt, bottom=5pt,
  fonttitle=\bfseries\small\color{gold},
  title={#1},
  beforeafter skip=8pt,
}

\newtcolorbox{coralbox}[1][]{
  enhanced, breakable,
  colback=corallight, colframe=coral,
  leftrule=4pt, toprule=0.6pt, bottomrule=0.6pt, rightrule=0.6pt,
  left=8pt, right=6pt, top=5pt, bottom=5pt,
  fonttitle=\bfseries\small\color{coral},
  title={#1},
  beforeafter skip=8pt,
}

\newtcolorbox{purplebox}[1][]{
  enhanced, breakable,
  colback=purplelight, colframe=purple,
  leftrule=4pt, toprule=0.6pt, bottomrule=0.6pt, rightrule=0.6pt,
  left=8pt, right=6pt, top=5pt, bottom=5pt,
  fonttitle=\bfseries\small\color{purple},
  title={#1},
  beforeafter skip=8pt,
}

\newtcolorbox{codebox}{
  enhanced, breakable,
  colback=slatelight, colframe=slate,
  boxrule=0.5pt,
  left=8pt, right=6pt, top=5pt, bottom=5pt,
  fontupper=\ttfamily\small\color{slate},
  beforeafter skip=8pt,
}

% Section divider band
\newcommand{\sectionband}[3]{%
  \begin{tcolorbox}[
    enhanced,
    colback=#3, colframe=#3,
    boxrule=0pt, arc=0pt,
    left=12pt, right=12pt, top=10pt, bottom=10pt,
    beforeafter skip=14pt,
  ]
    {\Large\bfseries\color{white}#1}\\[2pt]
    {\small\itshape\color{#3!30!white}#2}
  \end{tcolorbox}%
}

% ── List Styling ──────────────────────────────────────────
\setlist[itemize]{
  leftmargin=1.4em,
  itemsep=2pt,
  topsep=4pt,
  label={\color{teal}\textbullet}
}
\setlist[enumerate]{
  leftmargin=1.6em,
  itemsep=2pt,
  topsep=4pt,
}

% ── Table helpers ─────────────────────────────────────────
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

\newcommand{\theadrow}[1]{\rowcolor{#1}}
\newcommand{\altrow}{\rowcolor{rowalt}}

% ── Info pair row (two-col label/value) ───────────────────
\newcommand{\infopair}[4]{%
  \begin{tabular}{L{2.4cm} L{5.4cm} L{2.4cm} L{5.4cm}}
    \rowcolor{teallight}
    \textbf{\small\color{slate}#1} & \small #2 &
    \textbf{\small\color{slate}#3} & \small #4 \\
  \end{tabular}\par\vspace{3pt}%
}

% ═════════════════════════════════════════════════════════
\begin{document}
% ═════════════════════════════════════════════════════════

% ── COVER PAGE ────────────────────────────────────────────
\thispagestyle{empty}

\vspace*{3cm}

\begin{center}
  {\fontsize{52}{56}\selectfont\bfseries\color{teal}TIDE \& TABLE}\\[8pt]
  {\Large\itshape\color{gold}A Cozy Seafood Story}\\[30pt]

  \begin{tcolorbox}[
    enhanced,
    colback=teallight, colframe=teal,
    boxrule=1pt, arc=4pt,
    width=0.75\textwidth,
    halign=center,
    left=16pt, right=16pt, top=12pt, bottom=12pt,
  ]
    {\large\bfseries\color{slate}Product Requirements \& Design Document}\\[6pt]
    {\normalsize\color{darkgray}Version 1.0 \quad|\quad May 2026}\\[4pt]
    {\small\color{darkgray}Platform: Web (React + Canvas) \quad|\quad Single Player}
  \end{tcolorbox}

  \vspace{20pt}

  {\large\bfseries\color{slate}GAME PILLARS}\\[8pt]
  \begin{minipage}{0.82\textwidth}
    \begin{itemize}
      \item \textbf{Cozy \& Cinematic} --- warm pixel art with dramatic lighting and depth-of-field shading
      \item \textbf{Driven by Real Data} --- fish species, rarity, and habitat from the live GBIF Marine Life API
      \item \textbf{Story-Forward} --- a debt-ridden seafood store owner rebuilding her life and community
      \item \textbf{Four Interlocked Loops} --- Fish $\rightarrow$ Farm $\rightarrow$ Cook $\rightarrow$ Sell, each reinforcing the next
    \end{itemize}
  \end{minipage}

  \vspace{28pt}

  \begin{tabular}{C{3.6cm} C{3.6cm} C{3.6cm} C{3.6cm}}
    \rowcolor{lightgray}
    {\large\bfseries\color{teal}FISH} &
    {\large\bfseries\color{green}FARM} &
    {\large\bfseries\color{gold}COOK} &
    {\large\bfseries\color{coral}SELL} \\[6pt]
  \end{tabular}
\end{center}

\vfill
\begin{center}
  \small\itshape\color{darkgray}Confidential --- Internal Use Only
\end{center}

\newpage

% ── TABLE OF CONTENTS ─────────────────────────────────────
\tableofcontents
\newpage

% ═════════════════════════════════════════════════════════
\section{Executive Summary}
% ═════════════════════════════════════════════════════════

\textit{Tide \& Table} is a top-down, single-player, story-driven cozy game built as a browser-based React + Canvas application. The player steps into the worn boots of \textbf{Marina Calder}, a young woman who has inherited her grandmother's struggling seaside restaurant and must pay off a mounting debt while rediscovering her love for the ocean and the community around her.

The gameplay loop is divided into four interconnected scenes --- a Fishing Ocean, an Island Farm, a cozy Kitchen, and Marina's Storefront --- all connected by a persistent minimap. Every mechanic is grounded in real-world data: fish available at any location are pulled from the \textbf{GBIF Marine Life API} based on geographic coordinates, and recipes are powered by the \textbf{Spoonacular Recipe API} filtered to available ingredients.

\begin{tealbox}[Core Design Philosophy]
Every system should feel cozy, tactile, and alive. Fishing feels like real maritime discovery. Farming feels like patient tending. Cooking feels like craft. Selling feels like community. The data layer is invisible to the player --- it should feel like magic, not a database lookup.
\end{tealbox}

\vspace{8pt}
\begin{tabular}{L{2.4cm} L{5.4cm} L{2.4cm} L{5.4cm}}
\rowcolor{teallight}
\textbf{\small\color{slate}Genre}     & \small Cozy Simulation / Story RPG     & \textbf{\small\color{slate}Platform}  & \small Web Browser (React + Canvas) \\
\rowcolor{white}
\textbf{\small\color{slate}Perspective} & \small Top-Down 2.5D Pixel Art       & \textbf{\small\color{slate}Players}   & \small Single Player \\
\rowcolor{teallight}
\textbf{\small\color{slate}Core APIs} & \small GBIF Marine Life + Spoonacular  & \textbf{\small\color{slate}Runtime}   & \small 30--60 min story arc \\
\rowcolor{white}
\textbf{\small\color{slate}Visual}    & \small Pixel Art + Cinematic Lighting  & \textbf{\small\color{slate}Audience}  & \small Ages 14+, Cozy game fans \\
\end{tabular}

% ═════════════════════════════════════════════════════════
\section{Story \& Narrative}
% ═════════════════════════════════════════════════════════

\subsection{Premise}

Marina Calder returns to her grandmother's island after years away in the city. She finds Grandma Edith's beloved seafood restaurant --- \textit{The Saltwater Kitchen} --- boarded up and deep in debt. The bank gives Marina 30 in-game days to clear \$5,000 in debt or the property gets seized. Marina must fish the surrounding waters, cultivate the overgrown garden, cook Edith's legendary recipes, and charm the townsfolk back through the door.

\subsection{Story Arc}

\begin{enumerate}
  \item \textbf{Days 1--5 --- Arrival:} Tutorial phase. Marina discovers the tools, meets the island characters, learns the basics of each loop.
  \item \textbf{Days 6--15 --- Hustle:} Core gameplay unlocks. Rare fishing spots open. New crop seeds available. More NPCs visit the store.
  \item \textbf{Days 16--25 --- Discovery:} Marina finds her grandmother's lost recipe book. Legendary fish species become catchable. Debt urgency intensifies.
  \item \textbf{Days 26--30 --- Resolution:} Final push. Community rallies. Optional: clear debt early for a bonus ending. Story cinematic plays on success or failure.
\end{enumerate}

\subsection{Characters}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{2.8cm} L{2.4cm} L{3cm} L{5.4cm}}
\toprule
\rowcolor{slate}
\textbf{\color{white}Character} & \textbf{\color{white}Role} & \textbf{\color{white}Personality} & \textbf{\color{white}Gameplay Function} \\
\midrule
Marina Calder   & Player Character  & Determined, wry, sentimental & Main protagonist --- player-controlled \\
\altrow
Edith (Grandma) & Narrative Anchor  & Warm, wise, absent           & Guides via letters \& flashbacks \\
Theo            & Fisherman NPC     & Gruff but kind               & Unlocks fishing tips \& rare spots \\
\altrow
Lila            & Island Farmer     & Cheerful, resourceful        & Unlocks new seed varieties \\
The Banker      & Antagonist        & Cold, impatient              & Drives debt urgency \\
\altrow
Townsfolk NPCs  & Customers         & Varied archetypes            & Enter store, react to food, spend money \\
\bottomrule
\end{tabularx}

% ═════════════════════════════════════════════════════════
\section{Navigation System --- The Minimap}
% ═════════════════════════════════════════════════════════

\subsection{Overview}

A persistent minimap lives in the bottom-right corner of every scene. The map is a stylized, painterly top-down view of Marina's island and the surrounding ocean. It is always visible and can be clicked to expand to full screen.

\subsection{Map Layout}

\begin{goldbox}[Map Composition]
The island sits centered on the map and contains: the Storefront (top of island), the Kitchen (connected to storefront), and the Farm (behind the kitchen). The surrounding ocean is navigable fishing territory. Clickable zones pulse with a soft animated highlight to indicate they are interactive.
\end{goldbox}

\subsection{Navigation Rules}

\begin{itemize}
  \item Clicking the \textbf{Island} on the minimap transitions to the Store scene (or Farm / Kitchen depending on sub-zone tapped)
  \item Clicking the \textbf{Ocean} on the minimap transitions to the Fishing scene
  \item The minimap always shows a current scene indicator --- a small glowing dot on Marina's location
  \item Scene transitions use a cinematic fade + ambient sound crossfade (ocean waves fading in, crowd noise fading out)
  \item The minimap can be expanded to full screen --- in this mode, sub-zones are labeled and clickable
\end{itemize}

% ═════════════════════════════════════════════════════════
\newpage
\sectionband{PART I}{Core Game Loops \& Mechanics}{teal}
\section{Fishing Scene}
% ═════════════════════════════════════════════════════════

\subsection{Scene Overview}

The fishing scene takes place on the open ocean. The camera is top-down, slightly tilted for a 2.5D perspective. Marina's wooden boat floats in the center of the screen, gently bobbing on animated waves. The water is rendered with shader-style animated tiles that catch light dynamically --- deep blue in open water, turquoise near the coast.

\subsection{Location Selection --- The Pin Map}

Before entering the fishing scene for the first time each day, Marina is shown a pixel art world map. She drags and drops a pin anywhere on the map.

\begin{itemize}
  \item The pin automatically \textbf{snaps to the nearest coastal zone} with documented marine life data
  \item A tooltip appears showing the ocean region name (e.g., ``Gulf of Mexico --- Coastal Waters'') and a hint about what might be found there
  \item Geographic coordinates of the snapped pin are passed to the GBIF Marine Life API to fetch real species data for that region
  \item Different regions unlock across the story arc --- early game limits Marina to local waters; late game unlocks deep-sea spots
\end{itemize}

\begin{purplebox}[API Integration --- GBIF Marine Life]
\textbf{Endpoint:} \texttt{https://api.gbif.org/v1/occurrence/search?decimalLatitude=\{lat\}\&decimalLongitude=\{lng\}\&taxonKey=148}\\[4pt]
Response is parsed to extract: species names, IUCN conservation status (Endangered / Vulnerable / Least Concern), geographic range, and occurrence count. This data drives species rarity weights and flavor text in the game.
\end{purplebox}

\subsection{Fishing Mechanics}

\subsubsection{Boat Movement}
\begin{itemize}
  \item Marina's boat is the fixed anchor of the fishing scene --- it does not move
  \item Marina herself (as a small pixel character) can walk freely on the boat deck --- stern to bow
  \item Walking off the edge of the boat is blocked by invisible collision --- she cannot fall into the water
  \item A small idle animation plays when stationary: Marina pulls her hat down, looks at the horizon
\end{itemize}

\subsubsection{Casting}
\begin{itemize}
  \item The player \textbf{clicks anywhere on the water} to cast the fishing rod to that location
  \item A smooth arc animation plays --- the line sails out and lands with a small splash ripple effect
  \item After casting, the bobber floats on the surface, animated with gentle wave physics
  \item Bite timing is randomized within a species-dependent window (common fish: 2--8 sec; rare fish: 10--30 sec)
  \item A visual + audio cue triggers when a fish bites --- the bobber dips, water ripples expand, a soft chime plays
  \item The player must \textbf{click the bobber} during the bite window to begin the catch
\end{itemize}

\subsubsection{Catch Minigame}
\begin{itemize}
  \item A small tension bar appears --- green on the left, red on the right
  \item The player holds the mouse button to reel in; holding too hard (tension spikes red) risks snapping the line
  \item Different fish have different fight patterns: passive drift, aggressive zigzag, sudden burst
  \item Successfully landing the fish triggers a catch animation --- Marina lifts the rod, fish flops on deck
\end{itemize}

\subsection{Fish Rarity System}

Species rarity is derived from GBIF conservation data and occurrence frequency, mapped to a five-tier rarity system:

\vspace{4pt}
\begin{tabularx}{\textwidth}{C{1cm} L{2.2cm} L{3.8cm} L{2.4cm} L{2cm}}
\toprule
\rowcolor{teal}
\textbf{\color{white}Tier} & \textbf{\color{white}Label} & \textbf{\color{white}GBIF Basis} & \textbf{\color{white}Sell Modifier} & \textbf{\color{white}Catch Rate} \\
\midrule
1 & Common    & Occurrence $>$ 10,000       & 1.0$\times$ base & 60\% \\
\altrow
2 & Uncommon  & Occurrence 1,000--10,000    & 1.8$\times$ base & 25\% \\
3 & Rare      & Occurrence 100--1,000       & 3.5$\times$ base & 10\% \\
\altrow
4 & Legendary & Occurrence $<$ 100          & 7$\times$ base   & 4\%  \\
5 & Mythic    & IUCN Endangered / Critical  & Special (story)  & 1\%  \\
\bottomrule
\end{tabularx}

\vspace{6pt}
\begin{coralbox}[Design Note --- Mythic Fish]
Mythic fish (endangered species) are \textbf{never sold}. They are photograph-and-release story moments. Catching a Mythic triggers a short cinematic: Marina photographs the fish, writes in her journal, and releases it. This awards story XP and narrative dialogue --- not money. This is an intentional values statement about conservation.
\end{coralbox}

\subsection{Inventory (Boat Deck)}

\begin{itemize}
  \item Caught fish appear in a bottom-edge inventory strip --- pixel art fish icons with species name and rarity star
  \item Hovering a fish shows: species name, scientific name, origin region, rarity, estimated sell value
  \item The boat has a \textbf{12 fish slot} hold capacity per trip; a visual meter fills as slots are used
  \item ``Return to Island'' button glows in the top-right corner
\end{itemize}

% ═════════════════════════════════════════════════════════
\newpage
\sectionband{FARMING}{The Island Garden --- Grow What You Need}{green}
\section{Farming Scene}
% ═════════════════════════════════════════════════════════

\subsection{Scene Overview}

The farm is a top-down garden scene on the back of Marina's island. Warm afternoon light casts long pixel shadows across soil patches. The scene has a hand-drawn quality --- slightly imperfect tile borders, wildflowers between patches, a wooden fence with morning glories.

\subsection{Layout}

\begin{itemize}
  \item 12 soil patches arranged in a 4$\times$3 grid, each independently plantable
  \item A seed chest in the corner holds Marina's available seeds
  \item A water barrel --- interacting refills the watering can
  \item A compost bin that converts non-sellable plant matter into fertilizer
\end{itemize}

\subsection{Core Farming Loop}

\begin{enumerate}
  \item Walk over a soil patch to highlight it
  \item Press \texttt{[E]} or click the Seed button to open seed selection overlay
  \item Choose a seed from inventory --- seeds compatible with current unlocked recipes are shown prominently
  \item Seed is planted --- a seedling icon appears with a growth progress ring
  \item Each in-game day advances growth by one stage; watering speeds it up (3 stages/day watered vs.\ 1 for dry)
  \item When fully grown, walk over patch and press \texttt{[E]} to harvest
\end{enumerate}

\subsection{Plant Catalogue \& Growth Cycles}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{2.6cm} L{3cm} L{4.4cm} L{3.2cm}}
\toprule
\rowcolor{green}
\textbf{\color{white}Plant} & \textbf{\color{white}Growth (Dry/Watered)} & \textbf{\color{white}Recipes It Enables} & \textbf{\color{white}Notes} \\
\midrule
Garlic       & 4 / 2 days & Garlic Butter Scallops, Fish Soup  & Grows in clusters of 3 \\
\altrow
Dill         & 3 / 1 days & Salmon Dill Cream                  & Wilts if not harvested \\
Lemon Thyme  & 5 / 3 days & Herb-Crusted Halibut               & Perennial --- regrows after harvest \\
\altrow
Tomatoes     & 6 / 3 days & Cioppino, Marinara Prawns          & Requires 2 adjacent patches \\
Chili Pepper & 4 / 2 days & Spicy Tuna Tartare                 & Reduces NPC visit time \\
\altrow
Fennel       & 7 / 4 days & Roasted Branzino                   & Unlocked at Day 10 \\
Seaweed      & 2 / 1 days & Miso Clam Broth                    & Harvested from water edge, not soil \\
\bottomrule
\end{tabularx}

\subsection{Recipe-Driven Planting Suggestions}

The farming scene has a \textbf{Recipe Board} pinned to the fence. It shows the 3 most profitable recipes the player could currently make or is close to making, along with which ingredients they still need to grow. This bridges the farming and cooking loops without feeling like a task list.

% ═════════════════════════════════════════════════════════
\newpage
\sectionband{COOKING}{The Saltwater Kitchen --- Grandma's Recipes Live On}{gold}
\section{Cooking Scene}
% ═════════════════════════════════════════════════════════

\subsection{Scene Overview}

The kitchen is a top-down cozy interior. Warm lantern light, exposed brick, bundles of dried herbs on the ceiling. The countertop holds a chopping board, a stove, and a plating station. A large chalkboard on the wall lists today's specials. Ambient sounds: sizzling, gentle piano, the creak of the old building.

\subsection{The Recipe Book}

The Recipe Book is a physical-looking tome sitting on a wooden stand. Clicking it opens a full-screen overlay styled as a hand-illustrated cookbook.

\begin{itemize}
  \item Pages turn with a smooth flip animation and a paper rustle sound
  \item Each recipe occupies a spread --- left page has ingredients with checkmarks (green = in inventory, gray = missing), right page has illustrated dish + flavor text
  \item The book \textbf{only shows recipes where Marina has at least 60\% of required ingredients}
  \item Recipes are fetched from the Spoonacular API using available ingredient names as query parameters
  \item Recipes are filtered to seafood/fish categories and sorted by estimated sell value
\end{itemize}

\begin{purplebox}[API Integration --- Spoonacular Recipe API]
\textbf{Endpoint:} \texttt{https://api.spoonacular.com/recipes/findByIngredients?ingredients=\{list\}\&number=10\&ranking=2}\\[4pt]
\textbf{Response fields used:} title, missedIngredients, usedIngredients, nutrition (calories, protein), readyInMinutes.\\[2pt]
Sell price is computed client-side from rarity modifier + nutrition.protein + presentation tier.
\end{purplebox}

\subsection{Cooking Mechanics}

\subsubsection{Simple Cooking Flow}
\begin{enumerate}
  \item Select a recipe from the Recipe Book
  \item Required ingredients are automatically pulled from inventory and displayed on the countertop
  \item A 3-step cooking minigame plays: \textbf{Prep} (timed chopping), \textbf{Cook} (heat management), \textbf{Plate} (drag-to-arrange)
  \item Quality of performance affects dish presentation tier: Basic / Good / Excellent / Perfect
  \item Completed dish goes into the ``Ready to Serve'' queue (max 5 dishes at once)
\end{enumerate}

\subsubsection{Dish Pricing Formula}

\begin{tcolorbox}[
  colback=purplelight, colframe=purple,
  boxrule=0.5pt, left=8pt, right=8pt, top=6pt, bottom=6pt,
]
\textbf{\color{purple}Base Price = (Fish Rarity Modifier $\times$ \$8) + (Nutritional Score / 10) + (Presentation Tier Bonus)}\\[4pt]
\small\color{darkgray}\textit{e.g., Legendary fish (\$56) + Nutritional 82/10 (\$8.20) + Perfect tier (+\$4) = \$68.20 per dish}
\end{tcolorbox}

\subsection{Recipe Unlock Progression}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{1.8cm} L{3.8cm} L{3cm} L{1.8cm} L{2.8cm}}
\toprule
\rowcolor{gold}
\textbf{\color{white}Day} & \textbf{\color{white}Recipe} & \textbf{\color{white}Key Ingredient} & \textbf{\color{white}Difficulty} & \textbf{\color{white}Base Sell Value} \\
\midrule
Day 1  & Pan-Fried Whitefish         & Any Common Fish          & Easy   & \$12--18  \\
\altrow
Day 3  & Garlic Butter Scallops      & Scallops + Garlic        & Easy   & \$24--32  \\
Day 6  & Salmon Dill Cream Pasta     & Salmon + Dill            & Medium & \$38--50  \\
\altrow
Day 8  & Herb-Crusted Halibut        & Halibut + Lemon Thyme    & Medium & \$45--62  \\
Day 12 & Spicy Tuna Tartare          & Rare Tuna + Chili        & Hard   & \$58--80  \\
\altrow
Day 16 & Cioppino (Edith's Recipe)   & Mixed Rare Seafood       & Hard   & \$72--105 \\
Day 22 & Branzino en Papillote       & Legendary Fish + Fennel  & Expert & \$110--150 \\
\bottomrule
\end{tabularx}

% ═════════════════════════════════════════════════════════
\newpage
\sectionband{SELLING}{The Saltwater Kitchen --- Open for Business}{coral}
\section{Store Scene}
% ═════════════════════════════════════════════════════════

\subsection{Scene Overview}

The storefront is a top-down cozy restaurant interior --- four wooden tables with mismatched chairs, string lights hanging from exposed beams, a hand-lettered menu board. Marina stands behind the counter. The front door opens periodically as NPCs enter.

\subsection{The Debt System}

A horizontal progress bar dominates the top of the screen, styled as a weathered ledger page. The full bar represents \$5,000 debt. As money is earned, the debt bar fills from right to left with warm amber color.

\begin{itemize}
  \item The debt bar shows: current debt remaining, daily earnings, and days left
  \item A subtle anxiety effect activates when debt is above 80\%: the bar pulses red, ambient music becomes more tense
  \item A celebration effect plays each time a major milestone is cleared (25\%, 50\%, 75\%, 100\%)
  \item The daily minimum earning target is displayed as a thin marker on the bar --- miss it and the Banker sends a letter
\end{itemize}

\subsection{NPC Customer System}

\subsubsection{NPC Behavior Flow}
\begin{enumerate}
  \item NPCs spawn from the front door at randomized intervals (faster on busy days, story-driven surges)
  \item Each NPC walks to an available table and sits --- animated walk cycle with pixel art character
  \item A thought bubble appears showing a dish silhouette they're craving
  \item If the craved dish is available in the Ready-to-Serve queue, Marina auto-delivers it --- coins float up
  \item If the dish is \textbf{not} available, the NPC waits briefly, then leaves disappointed (no income)
  \item After eating, the NPC waves, leaves a coin pile on the table, and exits
\end{enumerate}

\subsubsection{NPC Types \& Modifiers}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{3cm} L{2.4cm} L{2.4cm} L{5.4cm}}
\toprule
\rowcolor{coral}
\textbf{\color{white}NPC Type} & \textbf{\color{white}Patience} & \textbf{\color{white}Tip Modifier} & \textbf{\color{white}Special Behavior} \\
\midrule
Regular Townsfolk & Medium (30 sec) & 1.0$\times$ & Returns frequently if satisfied \\
\altrow
Tourists          & Low (15 sec)    & 1.5$\times$ & Craves rare/exotic dishes \\
Fishermen         & High (60 sec)   & 0.8$\times$ & Will accept any seafood dish \\
\altrow
Critic (rare)     & Very Low (10 sec) & 3$\times$ & Triggers review event on success \\
Theo / Lila       & Infinite        & Special item & Triggers dialogue scenes \\
\bottomrule
\end{tabularx}

\subsection{Store Upgrades}

\begin{itemize}
  \item \textbf{New Tables} (+2 seats) --- allows more simultaneous customers
  \item \textbf{Decorations} (plants, art, new lighting) --- increases NPC patience by 10 sec
  \item \textbf{Better Kitchen Equipment} --- unlocks higher-tier cooking minigame bonuses
  \item \textbf{Chalkboard Specials} --- feature a dish for 2$\times$ visibility to incoming NPCs
\end{itemize}

% ═════════════════════════════════════════════════════════
\newpage
\sectionband{PART II}{Technical Architecture \& Data Integration}{slate}
\section{API Architecture}
% ═════════════════════════════════════════════════════════

\subsection{Marine Life --- GBIF API}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{3.6cm} X}
\toprule
\rowcolor{slate}
\textbf{\color{white}Field} & \textbf{\color{white}Detail} \\
\midrule
Base URL            & \texttt{https://api.gbif.org/v1/occurrence/search} \\
\altrow
Key Parameters      & \texttt{decimalLatitude}, \texttt{decimalLongitude}, \texttt{radius} (50 km), \texttt{classKey=186} (Actinopterygii) \\
Response Fields     & species, speciesKey, iucnRedListCategory, occurrenceCount, mediaType \\
\altrow
Rarity Mapping      & Occurrence count + IUCN status $\rightarrow$ Tier 1--5 as defined in Section 4.4 \\
Caching Strategy    & Results cached per coordinate grid cell (1$^\circ \times$ 1$^\circ$) for session duration \\
\altrow
Fallback            & Preset species pool per ocean region if API is unavailable \\
\bottomrule
\end{tabularx}

\subsection{Recipes --- Spoonacular API}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{3.6cm} X}
\toprule
\rowcolor{slate}
\textbf{\color{white}Field} & \textbf{\color{white}Detail} \\
\midrule
Base URL            & \texttt{https://api.spoonacular.com/recipes/findByIngredients} \\
\altrow
Key Parameters      & \texttt{ingredients} (comma-separated), \texttt{number=10}, \texttt{ranking=2} (minimize missing) \\
Response Fields     & title, missedIngredients, usedIngredients, nutrition, readyInMinutes \\
\altrow
Filtering           & Post-filter to seafood tag; exclude any recipe with 0 fish ingredients \\
Sell Price          & Computed client-side: rarity modifier + nutrition.protein + presentation tier \\
\altrow
Caching             & Cache recipes per ingredient combination hash; refresh on inventory change \\
\bottomrule
\end{tabularx}

\subsection{State Management}

The game state is managed in a single top-level React context with the following slices:

\begin{itemize}
  \item \texttt{player} --- position, scene, energy, name, day counter
  \item \texttt{inventory} --- fish[], herbs[], dishes[] each with id, species, rarity, quantity
  \item \texttt{debt} --- totalDebt, currentDebt, dailyEarnings, daysRemaining
  \item \texttt{farm} --- patches[12] each with: planted seed, growthStage, lastWatered, isReady
  \item \texttt{story} --- flags\{\}, characterMet\{\}, recipesUnlocked[], currentChapter
  \item \texttt{cache} --- apiResultsByCoord\{\}, recipesByIngredientHash\{\}
\end{itemize}

% ═════════════════════════════════════════════════════════
\newpage
\section{Visual Design System}
% ═════════════════════════════════════════════════════════

\subsection{Aesthetic Pillars}

\textit{Tide \& Table} sits at the intersection of classic 16-bit pixel art and modern cinematic presentation. The aesthetic uses the pixel art constraint to achieve a specific warmth and legibility, while layering modern techniques on top.

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{3.2cm} L{4.8cm} L{5.2cm}}
\toprule
\rowcolor{teal}
\textbf{\color{white}Pillar} & \textbf{\color{white}Technique} & \textbf{\color{white}Reference Mood} \\
\midrule
Cinematic Lighting  & Radial gradient light maps overlay pixel scene; dynamic day/night cycle & Studio Ghibli amber dusk \\
\altrow
Depth of Field      & Foreground elements blur softly; background has desaturated vignette & Stardew Valley meets old photograph \\
Water Rendering     & Animated sine-wave tile shader with light reflection spots & Moana meets SNES-era water \\
\altrow
Character Animation & 8-directional sprites, 4--6 frame cycles; exaggerated idle personality & Undertale charm, Celeste precision \\
UI Integration      & HUD elements styled as physical objects (ledger book, tackle box, cookbook) & Paper Mario physicality \\
\altrow
Color Temperature   & Scene-specific warm/cool shifts: kitchen is amber, ocean is cool teal & Visual emotional storytelling \\
\bottomrule
\end{tabularx}

\subsection{Scene Color Palettes}

\vspace{4pt}
\begin{tabular}{C{3.6cm} C{3.6cm} C{3.6cm} C{3.6cm}}
\cellcolor[HTML]{0D4F6B}\textbf{\color{white}FISHING} &
\cellcolor[HTML]{2E5B3A}\textbf{\color{white}FARMING} &
\cellcolor[HTML]{7A3B1E}\textbf{\color{white}KITCHEN} &
\cellcolor[HTML]{3D2B1F}\textbf{\color{white}STORE} \\[2pt]
\cellcolor[HTML]{0D4F6B}\small\color[HTML]{4BACC6}Accent: \#4BACC6 &
\cellcolor[HTML]{2E5B3A}\small\color[HTML]{7ABF73}Accent: \#7ABF73 &
\cellcolor[HTML]{7A3B1E}\small\color[HTML]{E8A860}Accent: \#E8A860 &
\cellcolor[HTML]{3D2B1F}\small\color[HTML]{D4956A}Accent: \#D4956A \\
\end{tabular}

\subsection{UI Component Style Guide}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{3.2cm} X}
\toprule
\rowcolor{slate}
\textbf{\color{white}Component} & \textbf{\color{white}Style Notes} \\
\midrule
Inventory Panels  & Worn wooden texture background, items as pixel icons with slot borders, tooltips on hover \\
\altrow
Debt Bar          & Parchment-style bar, ink-filling animation, quill cursor icon \\
Minimap           & Painted look, slight vignette at edges, pulsing zone indicators \\
\altrow
Dialogue Boxes    & Rounded-corner speech bubbles, character portrait left, text typewriter effect \\
Recipe Book       & Parchment pages with ink illustrations, handwritten-style font for recipe names \\
\altrow
Money Counter     & Brass coin icon, ticking counter animation on earn events, subtle jingle SFX \\
Buttons           & Carved-wood style with press-in depth effect, hover glow in scene accent color \\
\bottomrule
\end{tabularx}

\subsection{Lighting \& Shader Effects}

\begin{itemize}
  \item \textbf{Day/Night Cycle} --- 10 real-time minutes = 1 in-game day; light temperature shifts from cool morning blue $\rightarrow$ warm afternoon gold $\rightarrow$ dusky orange $\rightarrow$ night navy
  \item \textbf{Radial Light Sources} --- kitchen lanterns, store string lights, and boat lantern each emit soft radial gradients layered over the scene via Canvas compositing
  \item \textbf{Weather System} --- rainy days reduce fishing cast distance but increase rare fish probability; implemented via particle system + blue color shift overlay
  \item \textbf{Depth Vignette} --- a soft desaturated feathered overlay on scene edges pushes visual focus to the center of the playfield
\end{itemize}

% ═════════════════════════════════════════════════════════
\newpage
\section{Audio Design}
% ═════════════════════════════════════════════════════════

\subsection{Music}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{3.6cm} L{3.2cm} X}
\toprule
\rowcolor{teal}
\textbf{\color{white}Scene / State} & \textbf{\color{white}Mood} & \textbf{\color{white}Instrumentation} \\
\midrule
Fishing (Calm)    & Gentle, open      & Solo guitar + distant accordion + wave rhythm \\
\altrow
Fishing (Bite!)   & Brief tense sting & Pizzicato strings, quick 4-bar burst \\
Farming           & Pastoral, patient & Fingerpicked guitar + bird ambience + hand drum \\
\altrow
Kitchen (Cooking) & Lively, rhythmic  & Jazz piano + light percussion + sizzle texture \\
Store (Open)      & Cozy bustle       & Cafe guitar + soft chatter undertone \\
\altrow
Store (Rush)      & Elevated tempo    & Same motif, faster tempo, added bass \\
Night / Closing   & Reflective        & Solo piano, reverb-heavy, distant ocean \\
\altrow
Story Cinematics  & Orchestral swell  & Full arrangement, dynamic per story beat \\
\bottomrule
\end{tabularx}

\subsection{Sound Effects}

\begin{itemize}
  \item \textbf{Fishing cast:} rod whoosh + line hiss + splash
  \item \textbf{Fish bite:} bobber plunk + soft chime
  \item \textbf{Fish caught:} flop + water drip + success jingle
  \item \textbf{Crop harvest:} leaf rustle + pop + gentle chime
  \item \textbf{Dish completed:} sizzle fade + plate clink + coin shimmer
  \item \textbf{Customer satisfied:} happy hum + coin drop
  \item \textbf{Customer leaves unhappy:} disappointed sigh + door close
  \item \textbf{Debt milestone:} coin cascade + bell ring + crowd murmur
\end{itemize}

% ═════════════════════════════════════════════════════════
\section{Technical Architecture}
% ═════════════════════════════════════════════════════════

\subsection{Tech Stack}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{3.2cm} L{3.8cm} X}
\toprule
\rowcolor{slate}
\textbf{\color{white}Layer} & \textbf{\color{white}Technology} & \textbf{\color{white}Rationale} \\
\midrule
Rendering         & HTML5 Canvas + React          & Canvas for game scenes; React for UI overlays \\
\altrow
State Management  & React Context + useReducer    & Single source of truth; appropriate for game scale \\
Animation         & rAF loop + CSS transitions    & Game loop for scenes; CSS for UI transitions \\
\altrow
Pixel Art         & Tiled editor (JSON) + Canvas  & Efficient tile-based scene rendering \\
API Layer         & Custom fetch hooks + caching  & Resilient to rate limits; offline demo mode \\
\altrow
Marine Life API   & GBIF REST (free, no key)      & Real species data by coordinates \\
Recipe API        & Spoonacular REST (API key)    & Rich ingredient-based recipe search \\
\altrow
Storage           & localStorage                  & Persist progress across browser sessions \\
Hosting           & Vercel / Netlify              & Fast CDN; no backend required \\
\bottomrule
\end{tabularx}

\subsection{File Structure}

\begin{codebox}
src/\\
  game/           --- Canvas game engine (loop, scenes, entities)\\
    scenes/       --- FishingScene, FarmScene, KitchenScene, StoreScene\\
    entities/     --- Player, Fish, NPC, Crop, Dish\\
    systems/      --- PhysicsSystem, LightingSystem, ParticleSystem\\
  ui/             --- React overlay components\\
    Inventory.jsx, RecipeBook.jsx, DebtBar.jsx, Minimap.jsx\\
  api/            --- API integration hooks\\
    useMarineLife.js, useRecipes.js\\
  state/          --- GameContext, reducers, selectors\\
  assets/         --- sprites/, tilesets/, audio/, fonts/\\
  story/          --- dialogue trees, story flags, cutscenes
\end{codebox}

\subsection{Performance Targets}

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{3.2cm} L{2.4cm} X}
\toprule
\rowcolor{teal}
\textbf{\color{white}Metric} & \textbf{\color{white}Target} & \textbf{\color{white}Strategy} \\
\midrule
Frame Rate        & 60fps stable   & Dirty rectangle Canvas rendering; only redraw changed tiles \\
\altrow
Initial Load      & $<$ 4 seconds  & Sprite atlas compression; lazy-load non-active scenes \\
API Response      & $<$ 800ms perc.& Show loading animation; cache all responses \\
\altrow
Memory            & $<$ 150MB RAM  & Unload inactive scene assets; pool particle objects \\
Mobile            & Tablet touch   & Touch event layer over Canvas; responsive UI scaling \\
\bottomrule
\end{tabularx}

% ═════════════════════════════════════════════════════════
\newpage
\section{Development Roadmap}
% ═════════════════════════════════════════════════════════

\subsection*{Phase 1 --- Foundation (Weeks 1--2)}
\begin{itemize}
  \item Canvas game loop with scene switching and minimap navigation
  \item Pixel art tile renderer with fishing ocean tileset
  \item Basic player movement on boat deck with collision
  \item GBIF API integration with coordinate-to-species mapping and caching
  \item Inventory system with basic fish icons
\end{itemize}

\subsection*{Phase 2 --- Core Loops (Weeks 3--4)}
\begin{itemize}
  \item Fishing rod cast animation, bobber physics, catch minigame
  \item Farm scene with soil patches, planting, and growth cycle
  \item Spoonacular API integration with ingredient-to-recipe mapping
  \item Recipe Book UI with page-flip animation
  \item Kitchen cooking minigame (3 steps)
\end{itemize}

\subsection*{Phase 3 --- Selling \& Story (Weeks 5--6)}
\begin{itemize}
  \item Store scene with NPC spawn system and customer behavior AI
  \item Debt bar with milestone events
  \item Dialogue system with character portraits and typewriter text
  \item Day/night cycle with dynamic lighting layer
  \item Story chapter 1 (Days 1--10) dialogue and cutscenes
\end{itemize}

\subsection*{Phase 4 --- Polish (Weeks 7--8)}
\begin{itemize}
  \item Full audio implementation (music + SFX)
  \item Particle effects (water splash, coin burst, harvest pop)
  \item Weather system with rain particles
  \item Story chapters 2--4 and ending cinematics
  \item Store upgrade system and NPC type variety
  \item Performance optimization, save/load, mobile touch support
\end{itemize}

\subsection*{Reach Features}
\begin{itemize}
  \item 3D isometric rendering mode toggle (replace Canvas with Three.js)
  \item Hand gesture controls via MediaPipe (wave to cast, pinch to reel)
  \item Multiplayer co-op --- one player fishes, one farms, one cooks
  \item Seasonal events --- winter ice fishing, summer harvest festival
\end{itemize}

% ═════════════════════════════════════════════════════════
\section{Success Metrics}
% ═════════════════════════════════════════════════════════

\vspace{4pt}
\begin{tabularx}{\textwidth}{L{4cm} L{2.8cm} X}
\toprule
\rowcolor{teal}
\textbf{\color{white}Metric} & \textbf{\color{white}Target} & \textbf{\color{white}Measurement Method} \\
\midrule
Story Completion Rate     & $>$ 60\%        & Session tracking, save state analysis \\
\altrow
Session Length            & 25--40 min avg  & Time-in-scene logging \\
All Four Loops Played     & $>$ 80\%        & Scene visit flags \\
\altrow
API Integration Delight   & Qualitative     & Players notice fish are real species \\
Recipe Discovery          & 6+ recipes avg  & Recipe cook counter \\
\altrow
Debt Freedom              & $>$ 40\%        & Ending trigger events \\
\bottomrule
\end{tabularx}

% ═════════════════════════════════════════════════════════
\section{Appendix --- Open Questions \& Decisions}
% ═════════════════════════════════════════════════════════

\vspace{4pt}
\begin{tabularx}{\textwidth}{C{0.5cm} L{3.2cm} L{4cm} X}
\toprule
\rowcolor{slate}
\textbf{\color{white}\#} & \textbf{\color{white}Question} & \textbf{\color{white}Options} & \textbf{\color{white}Recommendation} \\
\midrule
1 & API Key for Spoonacular in demo? & Expose in client / proxy backend / mock data & Use a thin Vercel serverless function as proxy to protect the key \\
\altrow
2 & Save system scope? & localStorage / cloud save / no save & localStorage for demo; cloud save as Phase 5 reach \\
3 & Mobile-first or desktop-first? & Mobile (touch) / Desktop (mouse+kb) & Desktop primary; tablet touch as secondary \\
\altrow
4 & Minimap art style? & Pixel / painted / minimal schematic & Painted / impressionistic to match cozy aesthetic \\
5 & NPC generation? & Fixed set / procedural names & Fixed set for story; procedurally named townsfolk extras \\
\bottomrule
\end{tabularx}

% ── CLOSING ───────────────────────────────────────────────
\vspace{24pt}
\begin{center}
  \color{midgray}\rule{0.6\textwidth}{0.4pt}\\[6pt]
  \small\itshape\color{darkgray}Tide \& Table \ --- \ PRD v1.0 \ --- \ May 2026 \ --- \ Confidential
\end{center}

\end{document}